---
output:
  html_document: default
  pdf_document: default
---
Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:
If the preceived audience size perfectly match the real size, should follow the diagonal line
***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)
pf <- read.delim('pseudo_facebook.tsv')

ggplot(aes(x = age, y = friend_count), data = pf) + geom_point() 

summary(pf$age)
```

***

#### What are some things that you notice right away?
Response:
Young users have more friends. People lied ages or fake accounts. Strange high counts on specific old age.

***

### ggplot Syntax
Notes:
ggplot can help describe more complex data. The max age is 113, for people older than 90 the data are not very confident, should used limit on age.

```{r ggplot Syntax}
ggplot(aes(x = age, y = friend_count), data = pf) + geom_point() + xlim(13, 90)

summary(pf$age)
```

***

### Overplotting
Notes:
Overplotting makes it difficult to see actual number of points in each region.
Adding geom_jitter can make points more dispersive.

```{r Overplotting}
ggplot(aes(x = age, y = friend_count), data = pf) + geom_point(alpha = 1/20) + xlim(13, 90)

ggplot(aes(x = age, y = friend_count), data = pf) + geom_jitter(alpha = 1/20) + xlim(13, 90)

```

#### What do you notice in the plot?
Response:
Users under age 25 have more friend coutns. Age at about 70 have more friend counts, which is not consistent with other user ages.
***

### Coord_trans()
Notes:
Transformation on y axis can have a better vision of data.

```{r Coord_trans()}

```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
ggplot(aes(x=age, y=friend_count), data=pf) + 
  geom_jitter(alpha=1/20, position=position_jitter(h=0)) +
  xlim(13,90) +
  coord_trans(y='sqrt')
```

#### What do you notice?
If use geom_jitter, it's better to add position=position_jitter(h=0) since there will be negative values.
***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
ggplot(aes(x=age, y=friendships_initiated), data=pf) + 
  geom_jitter(alpha=1/10, position=position_jitter(h=0)) +
  xlim(13,90) +
  coord_trans(y='sqrt')

```

***

### Overplotting and Domain Knowledge
Notes:
Transform the x and y axis in precentage can give a better view of the relationship between actual audience and perceived audience.

***

### Conditional Means
Notes:
Scatter plot data stays very close to each other. It's better to do some calculation on data to have a better idea.

```{r Conditional Means}
library(dplyr)
age_groups <- group_by(pf, age)

pf.fc_by_age <- pf %>%
        group_by(age) %>%
        # summarise tells how many users in each group, n is number of users in each group
        summarise(friend_count_mean = mean(friend_count),
                  friend_count_median = median(friend_count),
                  n = n()) %>%
        arrange(age)
head(pf.fc_by_age)
```

Create your plot!

```{r Conditional Means Plot}
ggplot(aes(x=age, y=friend_count_mean), data=pf.fc_by_age) + geom_line()

```

***

### Overlaying Summaries with Raw Data
Notes:
The last plot could be used for quick summaries and exploration.

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x = age, y = friend_count), data = pf) + xlim(13, 90) + 
        coord_cartesian(xlim = c(13, 70), ylim = c(0, 1000)) + 
        geom_point(alpha = 0.05, 
                   position = position_jitter(h = 0),
                   color = 'orange') +
        # add mean to the plot
        geom_line(stat = 'summary', fun.y = mean) +
        # add 10% quantile to the plot
        geom_line(stat = 'summary', fun.y = quantile,
            fun.args = list(probs = .1), linetype = 3, color = 'blue' ) +
        # add 50% quantile to the plot
        geom_line(stat = 'summary', fun.y = quantile,
            fun.args = list(probs = .5), linetype = 3, color = 'blue' ) +        
        # add 90% quantile to the plot
        geom_line(stat = 'summary', fun.y = quantile,
            fun.args = list(probs = .9), linetype = 3, color = 'blue' )      

```

#### What are some of your observations of the plot?
Response:
Age around 30 to 70, has less than 250 friends on FB. People age below 30 have more friends and peak is above 750.
***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:
Histogram can give a better understanding of the summary and data based on what we have here. 
***

### Correlation
Notes:
User Pearons' R to find the linear relation between friend_count and age.
```{r Correlation}
?cor.test
cor.test(pf$age, pf$friend_count, method = 'pearson')

with(pf, cor.test(age, friend_count), method = 'pearson')
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:
Age and friend count is not linear.

***

### Correlation on Subsets
Notes:
Negative relationship from age and friend_count.

```{r Correlation on Subsets}
with(subset(pf, age <= 70), cor.test(age, friend_count, method = 'pearson'))
```

***

### Correlation Methods
Notes:
Monotonic relationship we can use spearman method.

```{r Correlation on Subsets}
with(subset(pf, age <= 70), cor.test(age, friend_count, method = 'spearman'))
```

***

## Create Scatterplots
Notes:

```{r}
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) + geom_point()

```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) + geom_point() +
        xlim(0, quantile(pf$www_likes_received, 0.95)) +
        ylim(0, quantile(pf$likes_received, 0.95)) +
        geom_smooth(method = 'lm', color = 'red')

```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
cor.test(pf$www_likes_received, pf$likes_received, method = 'pearson')

```

Response:
Strong positive correlation. 

***

### Moira on Correlation
Notes:


***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
# install.packages('alr3')
library(alr3)
data("Mitchell")
?Mitchell
```

Create your plot!

```{r Temp vs Month}
ggplot(aes(x = Temp, y = Month), data = Mitchell) + geom_point()

```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.
0

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
cor.test(Mitchell$Month, Mitchell$Temp, method = 'pearson')
```

***

### Making Sense of Data
Notes:
Data are in month, better to break by 12

```{r Making Sense of Data}
ggplot(aes(x = Month, y = Temp), data = Mitchell) + geom_point() +
        scale_x_discrete(breaks = seq(0, 203, 12)) 
```

***

### A New Perspective

What do you notice?
Response:
There's a sequence, like sine or cosine function

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:
Age mesured in month may give a better understanding of the data

```{r Understanding Noise: Age to Age Months}
pf$age_with_months <- with(pf, age + (1 - dob_month / 12))

```

***

### Age with Months Means

```{r Age with Months Means}
age_groups <- group_by(pf, age_with_months)
pf.fc_by_age_months <- summarise(age_groups,
                          friend_count_mean = mean(friend_count),
                          friend_count_median = median(friend_count),
                          n = n())
pf.fc_by_age_months <- arrange(pf.fc_by_age_months, age_with_months)
head(pf.fc_by_age)

```

Programming Assignment
```{r Programming Assignment}

```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
ggplot(aes(x = age_with_months, y = friend_count_mean), data = subset(pf.fc_by_age_months, age_with_months < 71)) +
        geom_line()

```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}
p1 <- ggplot(aes(x = age, y = friend_count_mean), data = subset(pf.fc_by_age, age < 71)) +
        geom_line() + geom_smooth()

p2 <- ggplot(aes(x = age_with_months, y = friend_count_mean), data = subset(pf.fc_by_age_months, age_with_months < 71)) +
        geom_line() + geom_smooth()

p3 <- ggplot(aes(x = round(age/5)*5, y = friend_count), data = subset(pf, age < 71)) +
        geom_line(stat = 'summary', fun.y = mean) 

library(gridExtra)
grid.arrange(p2, p1, p3, ncol = 1)

```

***

### Which Plot to Choose?
Notes:
Create multiple visulization to explore which is better

***

### Analyzing Two Variables
Reflection:
Scatter plot with conditional summary, correlation to understand relationship between two variables. How to use jitter and transparency to explore data.

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

