Lesson 5
========================================================

### Multivariate Data
Notes:
Add to color to represent the age in the scatter plot, but still hard to recognize any pattern.

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:
In the first graph, x is mark of the mean of the age. The mean age of female is bigger than male.
In the second graph, we can find the median friend count for female is larger than male for younger users, for older users it's hard to compare.

```{r Third Qualitative Variable}
library(ggplot2)
library(dplyr)
pf <- read.delim('pseudo_facebook.tsv')

ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + geom_boxplot() + 
        stat_summary(fun.y = mean, geom = 'point', shape = 4)

ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(gender))) + 
        geom_line(aes(color = gender), stat = 'summary', fun.y = median) 
```

```{r quiz}
pf.fc_by_age_gender <- subset(pf, !is.na(gender)) %>%
    group_by(age, gender) %>%
    summarise(mean_friend_count = mean(friend_count), 
              median_friend_count = median(friend_count), 
              n = n()) %>%
    ungroup() %>%
    arrange(age)
```
***

### Plotting Conditional Summaries
Notes:
use new date frame to create the same line graph as in last part. After we create new data frame, the plot code is shorter and more clear.

```{r Plotting Conditional Summaries}
ggplot(aes(x = age, y = mean_friend_count), data = pf.fc_by_age_gender) + geom_line(aes(color = gender))
```

***

### Thinking in Ratios
Notes:
This plot can help us examine the difference between male and female users.

***

### Wide and Long Format
Notes:
Reshape data to answer the question above. Convert long format to wide format. Add male and female column and remove the gender column to give a better view. dcast is for data frame and acast is for matrix or array. In front ~ means keep the data and after ~ means change the data to new column. 

***

### Reshaping Data
Notes:
Install new package to help us reshape data frame. 

```{r}
install.packages('reshape2')
library(reshape2)

pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender, 
                                  age ~ gender,
                                  value.var = 'median_friend_count')
head(pf.fc_by_age_gender.wide)
```


***

### Ratio Plot
Notes: 
Plot the female/male ratio. 
For young users, female have more friend counts. 

```{r Ratio Plot}
ggplot(aes(x = age, y = female/male), data = pf.fc_by_age_gender.wide) + geom_line() +
        geom_hline(yintercept = 1, alpha = 0.3, linetype = 2)
```

***

### Third Quantitative Variable
Notes:
Add year joined to the data frame, since young users have less time to acumulate more users. 

```{r Third Quantitative Variable}
pf$year_joined <- floor(2014 - pf$tenure/365)

```

***

### Cut a Variable
Notes:
Cut year joined into bucket to have a better understanding. 

```{r Cut a Variable}
summary(pf$year_joined)
table(pf$year_joined)

pf$year_joined.bucket <- cut(pf$year_joined, c(2004, 2009, 2011, 2012, 2014) )
```

***

### Plotting it All Together
Notes:
Create a line graph of friend_count vs. age so that each year_joined.bucket is a line tracking the median user friend_count across age. This means you should have four different lines on your plot.

```{r Plotting it All Together}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(year_joined.bucket))) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)
```

***

### Plot the Grand Mean
Notes:
From last plot, users joined earlier have more friend counts. Check bucket mean and grand mean also give us idea about what's data look like.

```{r Plot the Grand Mean}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(year_joined.bucket))) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean) + 
        geom_line(stat = 'summary', fun.y = mean, linetype = 2)

```

***

### Friending Rate
Notes:
We can find how many friends adding each day by looking at tenure

```{r Friending Rate}
with(subset(pf, tenure >= 1), summary(friend_count/tenure))

```

***

### Friendships Initiated
Notes:

What is the median friend rate?
0.2205
What is the maximum friend rate?
417.0000

```{r Friendships Initiated}
ggplot(aes(x = tenure, y = friendships_initiated/tenure), data = subset(pf, tenure >= 1)) + 
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean) 
```

***

### Bias-Variance Tradeoff Revisited
Notes:
With longger tenure, has less friendships

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
        geom_smooth(aes(color = year_joined.bucket))
```

***

### Sean's NFL Fan Sentiment Study
Notes:


***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
yo <- read.csv('yogurt.csv')
str(yo)

# change id from int to factor
yo$id <- factor(yo$id)
str(yo)

ggplot(aes(x = price), data = yo) + geom_histogram()
```

***

### Number of Purchases
Notes:

```{r Number of Purchases}
summary(yo)
yo <- transform(yo, all.purchases = strawberry + blueberry + pina.colada + plain + mixed.berry)

summary(yo$all.purchases)

```

***

### Prices over Time
Notes:

```{r Prices over Time}
ggplot(aes(x = time, y = price), data = yo) + geom_jitter(alpha = 1/4, shape = 21, fill = 'green')

```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
set.seed(4230)
sample.ids <- sample(levels(yo$id), 16)

ggplot(aes(x = time, y = price), data = subset(yo, id %in% sample.ids)) + 
        facet_wrap(~id) + 
        geom_line() + 
        geom_point(aes(size = all.purchases), pch = 1)
```

***

### The Limits of Cross Sectional Data
Notes:
If have observations over time, we can facet primary user case over time.

***

### Many Variables
Notes:


***

### Scatterplot Matrix
Notes:

```{r}
install.packages('GGally') 
library(GGally)
theme_set(theme_minimal(20))

set.seed(1836)
pf_subset <- pf[, c(2:15)]
names(pf_subset)
ggpairs(pf_subset[sample.int(nrow(pf_subset), 1000),])

```

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

