---
title: "lesson4_diamonds"
output: html_document
---

Lesson 4
========================================================

***

### Scatter plot price vs. x

```{r Scatterplots}
library(ggplot2)
data("diamonds")
names(diamonds)

ggplot(aes(x = x, y = price), data = diamonds) + geom_point() 

```

***

### Findings price vs. x
There are some data when x is 0 still has price values. Most of the x are in range 6-9 and price are also higher when x is 6-9

***

### Correlations
```{r Correlations}
cor.test(diamonds$x, diamonds$price, method = 'pearson')

cor.test(diamonds$y, diamonds$price, method = 'pearson')

cor.test(diamonds$z, diamonds$price, method = 'pearson')


```

***

### price vs. depth
```{r price vs. depth}
ggplot(aes(x = depth, y = price), data = diamonds) + geom_point() 

```

***

### Adjustments - price vs. depth
```{r Adjustments - price vs. depth}
ggplot(data = diamonds, aes(x = depth, y = price)) + 
        geom_point(alpha = 1/100) +
        scale_x_continuous(breaks = seq(quantile(diamonds$depth, 0.1), quantile(diamonds$depth, 0.9), 2))+
        xlim(quantile(diamonds$depth, 0.1), quantile(diamonds$depth, 0.9)) +
        ylim(quantile(diamonds$price, 0.1), quantile(diamonds$price, 0.9))       

```

***

### Correlation - price and depth
```{r Correlation - price and depth}
cor.test(diamonds$depth, diamonds$price, method = 'pearson')

```

The relationship between depth and price is weak

***

### price vs. volume
```{r price vs. volume}
diamonds$volume <- diamonds$x * diamonds$y *diamonds$z
count(diamonds$volume == 0)

ggplot(data = diamonds, aes(x = volume, y = price)) + 
        geom_point(alpha = 1/20) +
        scale_x_continuous(breaks = seq(quantile(diamonds$volume, 0.1), quantile(diamonds$volume, 0.9), 2))+
        xlim(quantile(diamonds$volume, 0.1), quantile(diamonds$volume, 0.9)) +
        ylim(quantile(diamonds$price, 0.1), quantile(diamonds$price, 0.9)) 

```

***

### Findings - price vs. volume
We can find the price and volume are in almost like a linear relation.

***

### Correlations on Subsets
```{r Correlations on Subsets}
with(subset(diamonds, volume <= 800 & volume >= 0), cor.test(volume, price, method = 'pearson'))

```

***

### Adjustments - price vs. volume
```{r Adjustments - price vs. volume}
ggplot(data = diamonds, aes(x = volume, y = price)) + 
        geom_point(alpha = 1/20) +
        scale_x_continuous(breaks = seq(quantile(diamonds$volume, 0.1), quantile(diamonds$volume, 0.9), 2))+
        xlim(0, 800) +
        ylim(quantile(diamonds$price, 0.1), quantile(diamonds$price, 0.9)) +
        geom_smooth(method = 'lm', color = 'red')

```

***

### Mean Price by Clarity
```{r Mean Price by Clarity}
suppressMessages(library(ggplot2))
suppressMessages(library(dplyr))
data(diamonds)

diamondsByClarity <- diamonds %>%
  group_by(clarity) %>%
  summarise(mean_price = mean(price), 
            median_price = median(price), 
            min_price = min(price), 
            max_price = max(price), 
            n= n())

diamondsByClarity

```

***

### Bar Charts of Mean Price
```{r Bar Charts of Mean Price}
diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))
head(diamonds_mp_by_clarity)

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))

p1  <- ggplot(diamonds_mp_by_clarity, aes(x = clarity, y = mean_price, fill= clarity)) +
  geom_bar(stat = "identity")

p2 <- ggplot(diamonds_mp_by_color, aes(x = color, y = mean_price, fill= color)) +
  geom_bar(stat = "identity")

library(gridExtra)
grid.arrange(p1, p2, ncol = 1)
```

***

### Trends in Mean Price
With color from D to J, the price is increasing. With clarity from l1 to IF, the price drops

***

### Gapminder Revisited
```{r Gapminder Revisited}
diamonds_by_cut <- group_by(diamonds, cut)
diamonds_mp_by_cut <- summarise(diamonds_by_cut, mean_price = mean(price))

ggplot(diamonds_mp_by_cut, aes(x = cut, y = mean_price, fill= cut)) +
  geom_bar(stat = "identity")
  
```


